Intentional release of native species undermines ecological stability

### 2022-03-30 rmd/manuscript.Rmd
edit: manuscript

---
title: "Intentional release of native species undermines ecological stability"
output:
  pdf_document:
    latex_engine: xelatex
  #word_document:
  # reference_docx: word_format.docx
csl: nature.csl
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile, encoding = encoding, output_dir = "document_output")
      })
bibliography: reference.bib
fontsize: 11pt
header-includes:
  - \usepackage{lineno}
  - \linenumbers
---

@@

The past decades have witnessed efforts to unveil ecological risks associated with massive releases of captive-bred native species ("intentional release"). Yet, this method is still pervasive in conservation and natural resource management, assuming that the demographic benefits exceed the associated risks. Here, contrary to this intuition, we show that intentional release triggers community collapse with little demographic benefit to the enhanced species. Our theory predicted that intentional release destabilizes community dynamics by facilitating competitive exclusion. In parallel, the excessive input of captive-bred individuals suppressed the natural recruitment of the enhanced species via intensified within-species competition. Consistent with this prediction, stream fish communities showed greater temporal fluctuations and fewer species richness in rivers with the intensive release of hatchery masu salmon - a major freshwater resource in Japan. Our findings alarm that the current overreliance on intentional release may accelerate global biodiversity loss, threatening the provisioning of ecosystem services.

# Maintext

@@

>>>>

---
title: "Intentional release of native species undermines ecological stability"
output:
  pdf_document:
    latex_engine: xelatex
  #word_document:
  # reference_docx: word_format.docx
csl: pnas.csl
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile, encoding = encoding, output_dir = "document_output")
      })
bibliography: reference.bib
fontsize: 11pt
header-includes:
  - \usepackage{lineno}
  - \linenumbers
---

@@

The past decades have witnessed efforts to unveil ecological risks associated with massive releases of captive-bred native species ("intentional release"). Yet, this method is still pervasive in conservation and natural resource management, assuming that the demographic benefits exceed the associated risks. However, such external inputs may disrupt the sensitive balance of species interactions that allows competing species to coexist, potentially compromising the long-term community stability. Here, we use theory and long-term data of stream fish communities to show that intentional release triggers community collapse with little demographic benefit to the enhanced species. Our theory predicted that intentional release intensifies interspecific competition, facilitating competitive exclusion of unenhanced species that otherwise stablely coexist. In parallel, the excessive input of captive-bred individuals suppressed the natural recruitment of the enhanced species via intensified within-species competition. Consequently, the ecological community with intentional release is predicted to show reduced community density with unstable temporal dynamics. Consistent with this prediction, stream fish communities showed greater temporal fluctuations and fewer species richness in rivers with the intensive release of hatchery salmon - a major fishery resource worldwide. Our findings alarm that the current overreliance on intentional release may accelerate global biodiversity loss, threatening the provisioning of ecosystem services.

# Significance

Intentional release of captive-bred individuals is a common practice for conservation and natural resource management. However, we know little about its potential consequences for the whole ecological community. Here, we show that intentional release undermines community stability with little demographic benefit to the enhanced species. Theory and data agreed that intentional release destabilizes community dynamics by facilitating competitive exclusion while suppressing the natural recruitment of the enhanced species. The effect size of intentional release was striking in its magnitude, doubling temporal fluctuations of enhanced communities compared to those with no intentional release. Our findings point to major limitations of intentional release as a primary tool for biodiversity conservation and sustainability.

**Key words:** coexistence theory, competition, resilience, ecological modeling, fishery

# Introduction

@@

# Results and Discussion
